\vspace{7cm} Equalizer 1.2\\Programming and User Guide

### 2012-02-27 Developer/ProgrammingGuide/paper.tex
1.2 PG work

\title{Equalizer\\Programming and User Guide}
\author{Eyescale Software GmbH}

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications using
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.10 for Equalizer 1.2\\[\medskipamount]
%  \today
}

@@

%\clearpage

@@

The images on the front page show the following applications build using
Equalizer: RTT DeltaGen\footnote{Image copyright Realtime Technology AG, 2008}
[top right], \textsf{Bino}, a stereo-capable, multi-display video
player\footnote{\link{http://bino.nongnu.org/}} [middle center], a flow
visualization application for climate research\footnote{Image courtesy of
  Computer Graphics and Multimedia Systems, University of Siegen} [middle
  right], the \textsf{eqPly} polygonal renderer in a three-sided CAVE [bottom
  left], the \textsf{eVolve} volume renderer\footnote{Data set courtesy of
  General Electric, USA} [bottom center], and the
\textsf{RTNeuron}\footnote{\link{http://bluebrain.epfl.ch/}} application to
  visualize cortical circuit simulations [bottom right].

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 09, 2012 & Revision for Equalizer 1.2\\
    \hline
  \end{tabularx}}
\clearpage

\pagenumbering{arabic}
\pagestyle{headings}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/tile.pdf}
  {\caption{\label{fPixel}\small A Tile Compound}}
\end{wrapfigure}
Tile compounds have a low, constant IO overhead for the image transfers and can
provide good scalability when used with fill-limited applications. Tile
decomposition works transparently for all Equalizer-based applications.

@@

{\footnotesize\lstinputlisting[firstline=57, lastline=68]
  {../../../Equalizer/examples/eqPly/main.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=80, lastline=88]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=90, lastline=100]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=101, lastline=118]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=132, lastline=143]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

@@

{\footnotesize\lstinputlisting[firstline=148, lastline=152]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=60, lastline=75]
  {../../../Equalizer/examples/eqPly/initData.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=122, lastline=129]
  {../../../Equalizer/examples/eqPly/frameData.h}}
{\footnotesize\lstinputlisting[firstline=54, lastline=80]
  {../../../Equalizer/examples/eqPly/frameData.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=143, lastline=174]
  {../../../Equalizer/examples/eqPly/vertexBufferDist.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=128, lastline=135]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=249, lastline=265B]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=588, lastline=625]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

\begin{wrapfigure}{r}{.618\textwidth}
%\begin{figure}[h!t]\center
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\label{fRender}Main Render Loop}}
\end{wrapfigure}
The model data is spatially organized in a 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree. During
this preprocessing step, each node of the tree gets a database range
assigned. The root node has the range [0, 1], its left child [0, 0.5]
and its right child [0.5, 1], and so on for all nodes in the tree. The
preprocessed model is saved in a binary format for accelerating
subsequent loading.

@@

Events are received by an event handler. The event handler finds the
\textsf{eq::Sys\-tem\-Win\-dow} associated to the event. It then creates a
generic \textsf{Event}, which holds the event data in an independent format. The
original native event and this generic \textsf{Event} form the
\textsf{SystemWindowEvent}, which is passed to the concrete
\textsf{SystemWindow} for processing.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
\textsf{Window::processEvent} is responsible for handling
the event locally and for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

For each GLX window, one \textsf{GLXEventHandler} is allocated. FBO's and
PBuffers are handled in the same way as window drawables. The event dispatch
finds the corresponding event handler for each received event, which is then
used to process the event in a fashion similar to AGL. The
\textsf{GLXWindowEvent} passes the \textsf{XEvent} to
\textsf{GLXWindowIF::\-process\-Event}.

@@

During this initialization, Equalizer creates and manages two threads
for communication, the receiver thread and the command thread. Normally
no application code is executed from these two threads.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\label{fThreads}Threads within one Node Process}}
\end{wrapfigure}
The receiver and command thread are terminated when the application
stops network communications using \textsf{Client::exitLocal}.

@@

{\footnotesize\lstinputlisting[firstline=70, lastline=106]
  {../../../Equalizer/libs/eq/client/glx/window.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=180, lastline=226]
  {../../../Equalizer/libs/eq/client/wgl/window.cpp}}

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \subfigure[]{\label{fFixed}
    \includegraphics[width=.382\textwidth]{images/focusFixed}}\hfil
  \subfigure[]{\label{fOrigin}
    \includegraphics[width=.382\textwidth]{images/focusOrigin}}\hfil
  \subfigure[]{\label{fObserver}
    \includegraphics[width=.382\textwidth]{images/focusObserver}}
  {\caption{Fixed(a) and dynamic focus distance relative to origin(b) and observer(c)}}
\end{wrapfigure}

@@

@@

\begin{wrapfigure}{r}{.612\textwidth}
  \includegraphics[width=.612\textwidth]{images/viewUML.pdf}
  {\caption{\label{fViewUML}UML Hierarchy of \textsf{eqPly::View}}}
\end{wrapfigure}
Layout API entities (Canvas, Segment, Layout, View) are sub-classed like all
other Equalizer entities using the \textsf{NodeFactory}. Equalizer registers the
master instance of these entities on the server node. Mutable parameters, e.g.,
the active layout of a canvas, are distributed using slave object commits
(cf. \sref{sSlaveCommit}). Application-specific data can be attached to a view
using a distributable \textsf{UserData} object. \fig{fViewUML} shows the UML
class hierarchy for the \textsf{eqPly::View}.

The render clients can access a slave instance of the view using
\textsf{Channel::getView}. When called from one of the frame task
methods, this method will return the view of the current destination
channel for which the task method is executed. Otherwise it returns the
channel's native view, if it has one. Only destination channels of an
active canvas have a native view.

@@

The default Equalizer event handling is using the view API to maintain
the aspect ratio of destination channels after a window resize. Without
updating the wall or projection description, the rendering would become
distorted.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
When a window is resized, a \textsf{CHANNEL\_RESIZE} event is eventually
generated. If the corresponding channel has a view,
\textsf{Channel::processEvent} sends a \textsf{VIEW\_RESIZE} event to
the application. This event contains the identifier of the view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\label{fDirectSend}Direct Send Compositing}}
\end{wrapfigure}
To provide a motivation for the design of the image compositing
API, the direct send parallel compositing algorithm is introduced in this
section. Other parallel compositing algorithms, e.g. binary-swap, can
also be expressed through an Equalizer configuration file.

@@

In direct send compositing, each rendering channel is also responsible
for the sort-last composition of one screen-space tile. He receives
the framebuffer pixels for his tile from all the other channels. The
size of one tile decreases linearly with the number of source channels,
which keeps the total amount of pixel data per channel constant.

@@

Readback and assemble operations on frames and images are designed to be
asynchronous. They have a start and finish method for both readback and
assemble to allow the initiation and synchronization of the operation.
Currently, only synchronous readback and assembly using
\textsf{glReadPixels} and \textsf{glDrawPixels} is implemented in the
respective start method of the image. Later versions of Equalizer will
implement asynchronous pixel transfers.

@@

\begin{wrapfigure}{r}{.618\textwidth}
%\begin{figure}[h!t]\center
  \includegraphics[width=.618\textwidth]{images/compositor.pdf}
  {\caption{\label{fCompositor}Functional Diagram of the Compositor}}
\end{wrapfigure}
On a high level, the compositor combines multiple input frames using 2D
tiling, depth-compositing for polygonal data or sorted, alpha-blended
compositing for semitransparent volumetric data. These operations
composite either directly all images on the GPU, or use a CPU-based
compositor and then transfer the preintegrated result to the GPU. The
high-level entry points automatically select the best algorithm. The
CPU-based compositor uses, based on availability, OpenMP and the
Paracomp compositing library to accelerate its operation.

@@

{\footnotesize\lstinputlisting[firstline=317, lastline=344]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=544, lastline=567]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/statistics}
  {\caption{\label{fStatistics}Statistics for a two node 2D compound}}
\end{wrapfigure}
Statistics are measured in milliseconds since the configuration was
initialized. The server synchronizes the per-configuration clock on each node
automatically. Each statistic event records the originator's (channel, window,
pipe, node, view or config) unique identifier.

Statistics are enabled per entity using an attribute hint. The hint determines
how precise the gathered statistics are. When set to \textsf{fastest}, the
per-frame clock is sampled directly when the event occurs. When set to
\textsf{nicest}, all OpenGL commands will be finished before sampling the
event. This may incur a performance penalty, but gives more correct results. The
default setting is fastest in release builds, and nicest in debug builds. The
fastest setting often attributes times to the operation causing an OpenGL
synchronization instead of the operation submitting the OpenGL commands, e.g.,
the readback time contains operations from the preceding draw operation.

@@

{\footnotesize\lstinputlisting[firstline=187, lastline=208]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

\cleardoublepage
\pagestyle{empty}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaserBack.pdf}
\end{textblock}

>>>>

\title{\vspace{7cm}
Equalizer 1.2\\Programming and User Guide}
\author{Eyescale Software GmbH}

%\date{
%  \vspace{14cm}
%  \begin{textblock}{}(0,4.5)
%    \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
%  \end{textblock}
%      {\Large The official reference for developing and
%        deploying parallel, scalable OpenGL\texttrademark\ applications using
%        the Equalizer parallel rendering framework}\\\vspace{1cm}
%      Version 1.10 for Equalizer 1.2\\[\medskipamount]
%  \today
%}

@@

\clearpage

@@

The images on the front page show the following applications build using
Equalizer: RTT DeltaGen\footnote{Image copyright Realtime Technology AG, 2008}
[top right], \textsf{Bino}, a stereo-capable, multi-display video
player\footnote{\link{http://bino.nongnu.org/}} [middle center], a flow
visualization application for climate research\footnote{Image courtesy of
  Computer Graphics and Multimedia Systems, University of Siegen} [middle
  right], the \textsf{eqPly} polygonal renderer in a three-sided CAVE [bottom
  left], the \textsf{eVolve} volume renderer\footnote{Data set courtesy of
  General Electric, USA} [bottom center], and the
\textsf{RTNeuron}\footnote{Image courtesy of
  \htmladdnormallink{Cajal Blue Brain}{http://cajalbbp.cesvima.upm.es/} /
  \htmladdnormallink{Blue Brain Project}{http://bluebrain.epfl.ch}}
  application to visualize cortical circuit simulations [bottom right].

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 17, 2012 & Revision for Equalizer 1.2\\
    \hline
  \end{tabularx}}
\clearpage

\pagenumbering{arabic}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/tile.pdf}
  {\caption{\label{fTile}\small A Tile Compound}}
\end{wrapfigure}
Tile compounds have a low, constant IO overhead for the image transfers and can
provide good scalability when used with fill-limited applications. Tile
decomposition works transparently for all Equalizer-based applications.

@@

{\footnotesize\lstinputlisting[firstline=58, lastline=68]
  {../../../Equalizer/examples/eqPly/main.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=81, lastline=89]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=91, lastline=101]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=102, lastline=119]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=134, lastline=143]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=145, lastline=159]
  {../../../Equalizer/examples/eqPly/eqPly.cpp}}

@@

@@

{\footnotesize\lstinputlisting[firstline=65, lastline=76]
  {../../../Equalizer/examples/eqPly/initData.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=123, lastline=130]
  {../../../Equalizer/examples/eqPly/frameData.h}}
{\footnotesize\lstinputlisting[firstline=55, lastline=84]
  {../../../Equalizer/examples/eqPly/frameData.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=142, lastline=174]
  {../../../Equalizer/examples/eqPly/vertexBufferDist.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=127, lastline=135]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=249, lastline=265]
  {../../../Equalizer/examples/eqPly/config.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=584, lastline=621]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[h!t]\center
  \includegraphics[width=.9\textwidth]{images/render.pdf}
  {\caption{\label{fRender}Main Render Loop}}
\end{figure}
The model data is spatially organized in a 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree. During
this preprocessing step, each node of the tree gets a database range
assigned. The root node has the range [0, 1], its left child [0, 0.5]
and its right child [0.5, 1], and so on for all nodes in the tree. The
preprocessed model is saved in a binary format for accelerating
subsequent loading.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-2ex}\includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\label{fEventProcessing}Event Processing}}\vspace{-4ex}
\end{wrapfigure}
Events are received by an event handler. The event handler finds the
\textsf{eq::Sys\-tem\-Win\-dow} associated to the event. It then creates a
generic \textsf{Event}, which holds the event data in an independent format. The
original native event and this generic \textsf{Event} form the
\textsf{SystemWindowEvent}, which is passed to the concrete
\textsf{SystemWindow} for processing.

@@

\textsf{Window::processEvent} is responsible for handling
the event locally and for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/eventUML.pdf}
  {\caption{\label{fEventUML}UML Class Diagram for Event Handling}}
\end{wrapfigure}
For each GLX window, one \textsf{GLXEventHandler} is allocated. FBO's and
PBuffers are handled in the same way as window drawables. The event dispatch
finds the corresponding event handler for each received event, which is then
used to process the event in a fashion similar to AGL. The
\textsf{GLXWindowEvent} passes the \textsf{XEvent} to
\textsf{GLXWindowIF::\-process\-Event}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\label{fThreads}Threads within one Node Process}}
\end{wrapfigure}
During this initialization, Equalizer creates and manages two threads
for communication, the receiver thread and the command thread. Normally
no application code is executed from these two threads.

@@

The receiver and command thread are terminated when the application
stops network communications using \textsf{Client::exitLocal}.

@@

{\footnotesize\lstinputlisting[firstline=70, lastline=104]
  {../../../Equalizer/libs/eq/client/glx/window.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=180, lastline=224]
  {../../../Equalizer/libs/eq/client/wgl/window.cpp}}

@@

@@

\begin{figure}[h!t]\center
%\begin{wrapfigure}{r}{.382\textwidth}
  \subfigure[]{\label{fFixed}
    \includegraphics[width=.32\textwidth]{images/focusFixed}}\hfil
  \subfigure[]{\label{fOrigin}
    \includegraphics[width=.32\textwidth]{images/focusOrigin}}\hfil
  \subfigure[]{\label{fObserver}
    \includegraphics[width=.32\textwidth]{images/focusObserver}}
  {\caption{Fixed(a) and dynamic focus distance relative to origin(b) and observer(c)}}
\end{figure}

@@

Layout API entities (Canvas, Segment, Layout, View) are sub-classed like all
other Equalizer entities using the \textsf{NodeFactory}. Equalizer registers the
master instance of these entities on the server node. Mutable parameters, e.g.,
the active layout of a canvas, are distributed using slave object commits
(cf. \sref{sSlaveCommit}). Application-specific data can be attached to a view
using a distributable \textsf{UserData} object. \fig{fViewUML} shows the UML
class hierarchy for the \textsf{eqPly::View}.

\begin{wrapfigure}{r}{.612\textwidth}
  \includegraphics[width=.612\textwidth]{images/viewUML.pdf}
  {\caption{\label{fViewUML}UML Hierarchy of \textsf{eqPly::View}}}
\end{wrapfigure}
The render clients can access a slave instance of the view using
\textsf{Channel::getView}. When called from one of the frame task
methods, this method will return the view of the current destination
channel for which the task method is executed. Otherwise it returns the
channel's native view, if it has one. Only destination channels of an
active canvas have a native view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
The default Equalizer event handling is using the view API to maintain
the aspect ratio of destination channels after a window resize. Without
updating the wall or projection description, the rendering would become
distorted.

When a window is resized, a \textsf{CHANNEL\_RESIZE} event is eventually
generated. If the corresponding channel has a view,
\textsf{Channel::processEvent} sends a \textsf{VIEW\_RESIZE} event to
the application. This event contains the identifier of the view.

@@

To provide a motivation for the design of the image compositing
API, the direct send parallel compositing algorithm is introduced in this
section. Other parallel compositing algorithms, e.g. binary-swap, can
also be expressed through an Equalizer configuration file.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\label{fDirectSend}Direct Send Compositing}}
\end{wrapfigure}
In direct send compositing, each rendering channel is also responsible
for the sort-last composition of one screen-space tile. He receives
the framebuffer pixels for his tile from all the other channels. The
size of one tile decreases linearly with the number of source channels,
which keeps the total amount of pixel data per channel constant.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\label{fAssembly}Hierarchy of Assembly Classes}}
\end{wrapfigure}
Readback and assemble operations on frames and images are designed to be
asynchronous. They have a start and finish method for both readback and
assemble to allow the initiation and synchronization of the operation.
Currently, only synchronous readback and assembly using
\textsf{glReadPixels} and \textsf{glDrawPixels} is implemented in the
respective start method of the image. Later versions of Equalizer will
implement asynchronous pixel transfers.

@@

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[h!t]\center
  \includegraphics[width=.9\textwidth]{images/compositor.pdf}
  {\caption{\label{fCompositor}Functional Diagram of the Compositor}}
\end{figure}
On a high level, the compositor combines multiple input frames using 2D
tiling, depth-compositing for polygonal data or sorted, alpha-blended
compositing for semitransparent volumetric data. These operations
composite either directly all images on the GPU, or use a CPU-based
compositor and then transfer the preintegrated result to the GPU. The
high-level entry points automatically select the best algorithm. The
CPU-based compositor uses, based on availability, OpenMP and the
Paracomp compositing library to accelerate its operation.

@@

{\footnotesize\lstinputlisting[firstline=320, lastline=344]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

{\footnotesize\lstinputlisting[firstline=542, lastline=563]
  {../../../Equalizer/examples/eqPly/channel.cpp}}

@@

Statistics are measured in milliseconds since the configuration was
initialized. The server synchronizes the per-configuration clock on each node
automatically. Each statistic event records the originator's (channel, window,
pipe, node, view or config) unique identifier.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/statistics}
  {\caption{\label{fStatistics}Statistics for a two node 2D compound}}
\end{wrapfigure}
Statistics are enabled per entity using an attribute hint. The hint determines
how precise the gathered statistics are. When set to \textsf{fastest}, the
per-frame clock is sampled directly when the event occurs. When set to
\textsf{nicest}, all OpenGL commands will be finished before sampling the
event. This may incur a performance penalty, but gives more correct results. The
default setting is fastest in release builds, and nicest in debug builds. The
fastest setting often attributes times to the operation causing an OpenGL
synchronization instead of the operation submitting the OpenGL commands, e.g.,
the readback time contains operations from the preceding draw operation.

@@

{\footnotesize\lstinputlisting[firstline=188, lastline=209]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

%\cleardoublepage
%\pagestyle{empty}
%\begin{textblock}{}(0,4.5)
%  \hspace{-1cm}\includegraphics[width=22cm]{images/teaserBack.pdf}
%\end{textblock}
